---
output:
  word_document: default
  html_document: default
---
# Module 2 - Assignment 2
## MacBride, Erica
### Introduction to R Markdown  
This is the first R Markdown file I have created. This file will go over topics we have discussed in class. The topic include vectors, data frames and data types. We have learned that the most commonly used data types are Logical, Character, and Doubles.  
The first chunk of code that I will be entering is a basic plot. The plot will show sales by month. First we must install the tidyverse package. 

```{r Plot of Sales by Month}
library(tidyverse)
sales <- c(50, 150, 200, 100, 300, 250)
month <- c(1, 2, 3, 4, 5, 6)
qplot(month, sales)
```

May had the largest amount of sales with a total of 300.  

```{r WWrking with Data Frames}
Yearly_Sales <- data.frame(month=c("Jan","Feb","Mar", "Apr","May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"), sales=c(150.25, 258.54,268.55, 122.52, 987, 458.82, 667.23, 845.54, 586.78, 888.58, 756.12, 456.84))
```

May had the largest amount of sales. The total sales for the month of may were 987. The month with the least amount of sales was April, with a total of 122.52.