---
output:
  word_document: default
  html_document: default
---
# Module 3 - Assignment 1
## MacBride, Erica
### Data Visualization

For this assignment we will be using the candy rankings data set and the candy production data set which can be found on the Canvas Course website under Module 3, Assignment 1.  

```{r}
library(tidyverse)

candy_data <- read_csv("candy_data.csv")
candy_production <- read_csv("candy_production.csv", 
    col_types = cols(observation_date = col_date(format = "%m/%d/%Y")))
```

#### Visualization with Scatterplots (geom_point)
```{r Sugar by Price Scatterplot}
ggplot(data=candy_data, aes(x=sugarpercent,y=pricepercent, label = competitorname)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(check_overlap = T, vjust = "bottom",nudge_y = 0.01, angle = 30, size = 2 ) +
  labs(title = "Sugar by Price Scatterplot", x = "Sugar Content (Percentile)", y = "Price (Percentile)")
```

Skittles original has the most sugar and lowest price. The Milky Way Simply Caramel has the highest sugar content and is also the most expensive.  

#### Line Chart of Candy Production
I will be using the Candy Production dataset which will display a specific date and how production during that month is compared to 2012.  

```{r Line Plots for Production}
ggplot(data=candy_production, aes(x=observation_date,y=IPG3113N)) +
  geom_line() +
  geom_smooth() +
  labs(title= "Monthly Candy Production", x= "Year", y= "As Percent of 2012 Production")
```

#### Bar chart of Ingredients  

```{r}
ggplot(data=candy_data, aes(x=chocolate)) +
  geom_bar()
candyFeatures <- candy_data %>% select(2:10)
candyFeatures[] <- lapply(candyFeatures, as.logical)

ggplot(data=candyFeatures, aes(x=chocolate)) +
  geom_bar()
```

