---
output:
  word_document: default
  html_document: default
---
# Module 5 - Assignment 1
## MacBride, Erica
### Data Wrangling

```{r}
library(tidyverse)
library(readxl)
UN_migrant <- read_excel("UN_migrant.xlsx", 
    sheet = "Table 6", col_types = c("numeric", 
        "text", "text", "numeric", "text", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric"), skip = 15)
```

### Part 2 - Cleaning Data with dplyr
```{r Renaming Columns}
UN_migrant<- rename(UN_migrant, Country = "...2", Country_Code = "...4", Type = "...5", "1990" = "1990...6", "1995" = "1995...7", "2000" = "2000...8", "2005" = "2005...9", "2010" = "2010...10", "2015" = "2015...11")
```

```{r New Subset}
Migration <- select(UN_migrant, Country, Country_Code, Type, "1990", "1995", "2000", "2005", "2010", "2015")
```

### Creating Tidy data using tidyr
```{r Migration Tidy}
Migration2 <- pivot_longer(Migration, c("1990", "1995", "2000", "2005", "2010", "2015"), names_to = "Year",values_to = "Cases" )
Migration2$Year <- as.integer(Migration2$Year)
```

### Part 4 - Research Questions
```{r New Subsets}
RegionalMigration <- Migration2 %>%
  filter(Country %in% c("Africa", "Asia", "Europe","Oceania", "Northern America", "South America"))

Americas <- Migration2 %>% 
  filter(Country %in% c("Central America", "South America", "Northern America"))
```

### Worldwide Migration based on Regions
Which region in the world has the highest number of migrants in the year 2005? Asia.
Over the years, which region consistently has the most migrants every 5 year span? Which has the second most?

```{r}
ggplot(data=RegionalMigration, aes(x=Year, y=Cases, color = Country)) +
  geom_point() +
  labs(title = "Worldwide Migration",
       x = "Year",
       y = "Migrants")

ggplot(data=RegionalMigration, aes(x=Year, y=Cases, fill = Country)) +
  geom_col() +
  scale_fill_discrete(name= "Region of World") +
  labs(title = "Worldwide Migration",
       x = "Year",
       y = "Migrants")
```
Which region in the world has the highest number of migrants in the year 2005? 
    Asia.

Over the years, which region consistently has the most migrants every 5 year span? Which has the second most?
    Asia has the most migrants every 5 years. Africa has the second most. 

What region has seen the fewest migrants over the years?
    Oceania.

Which plot was most useful in answering these questions and why?
    The bar plot with fill was most useful in answering these questions, as the fill illustrates a better idea of how      many migrants vs. the scatterplot.

### Americas Migration by Region

```{r}
ggplot(data=Americas, aes(x=Year, y=Cases, color = Country)) +
  geom_point() +
  labs(title = "Worldwide Migration",
       x = "Year",
       y = "Migrants")

ggplot(data=Americas, aes(x=Year, y=Cases, fill = Country)) +
  geom_col() +
  scale_fill_discrete(name= "Americas Region") +
  labs(title = "Worldwide Migration",
       x = "Year",
       y = "Migrants")
```

a. In 1990, which region had the largest number of migrants for the Americas?
    Central America.

b. Has this region continued to grow since 1990?
    Northern America. 

c. What trends do you notice happening in the Americas over the years?
    Central America saw a large increase in migrants in 1990 but then experience a significant decrease. Northern          America has seen a a steady trend in migrants from 1990 to 2015 although it is decreasing. South America follows a     a steady pattern of migrants increasing and decreasing around the same levels until 2010-2015 with a rapid             increase.
    
d. Specifically, has Northern America increased or decreased over the years?
    Decreased since 1995.
    
e. Which plot was most useful in answering these questions and why?
    The scatterplot as it shows trends much easier. 
    