---
output:
  word_document: default
  html_document: default
---
# Module 6 - Assignment 1
## MacBride, Erica
### Lubridate

```{r Loading Packages, message=FALSE}
library("tidyverse")
library("lubridate")
Appointments <- read_csv("Appointments.csv")
```

### Working with Dates
```{r Creating Reservation Year}
Appointments <- mutate(Appointments, "ReservationYear" = 2019 )
Appointments <- mutate(Appointments, "ReservationDate"= make_date(month=ReservationMonth,day=ReservationDay,year=ReservationYear))
Appointments <- mutate(Appointments, "CreationDate" = make_date(month=CreationMonth, day=CreationDay, year=CreationYear))
Appointments <- mutate(Appointments, "ReservationSpan" = ReservationDate - CreationDate)
```

### Exploratory Analysis
```{r}
summary(Appointments)
Appointments$ReservationSpan <- as.numeric(Appointments$ReservationSpan)
summary(Appointments)
```
a. What is the average time between appointment creation and the actual reservation date?
  38.53 days

b. What was the max time?
  207 days

c. What was the min time?
  1 day
  
```{r}
cor(Appointments$ReservationSpan, Appointments$Show)
```
  a. Are ReservationSpan and Show highly correlated?
      No, reservation span and show are not highly correlated.  

```{r}
ggplot(Appointments, aes(Show))+
  geom_bar()
```
a. Based on 0 being “no show” and 1 being “show”, did most people make their appointments?
  Yes, most people made their appointments.

