---
output:
  word_document: default
  html_document: default
---
# MIS503 - Final Project
## MacBride, Erica
### Zillow Home Value Index Analysis

```{r}
library("tidyverse")
SingleFamilyResidenceSales <- read_csv("SingleFamilyResidenceSales.csv")
SingleFamilyResidenceRental <- read_csv("SingleFamilyResidenceRental.csv")
```

### Wake County Home Sales
```{r}
WakeCountySales <- select(SingleFamilyResidenceSales, RegionName, State, CountyName, Metro, "1996-05", "1997-05", "1998-05","1999-05", "2000-05", "2001-05", "2002-05","2003-05", "2004-05", "2005-05","2006-05","2007-05", "2008-05","2009-05","2010-05", "2011-05","2012-05","2013-05","2014-05", "2015-05","2016-05","2017-05","2018-05")
WakeCountySales <- filter(WakeCountySales, State =="NC", CountyName == "Wake County")
WakeCountySales <- rename(WakeCountySales,"1996"="1996-05","1997"="1997-05","1998"="1998-05","1999"="1999-05","2000"="2000-05", "2001"="2001-05","2002"="2002-05","2003"="2003-05","2004"="2004-05","2005"="2005-05","2006"="2006-05","2007" ="2007-05","2008"= "2008-05","2009"= "2009-05","2010"= "2010-05","2011"="2011-05","2012" = "2012-05","2013"= "2013-05","2014"="2014-05", "2015"="2015-05","2016"= "2016-05","2017"="2017-05","2018"="2018-05" )
WakeCountySales<-pivot_longer(WakeCountySales, c("1996","1997","1998","1999","2000","2001","2002","2003","2004","2005","2006","2007", "2008","2009","2010","2011","2012","2013","2014","2015","2016","2017","2018"), names_to = 'YR',values_to = 'ZHVI')
```

```{r}
ggplot(data=WakeCountySales, aes(x=YR, y=ZHVI, color = RegionName)) +
  geom_point() +
  labs(title = "Wake County Home Sales",
       x = "Year",
       y = "Home Value Index")+
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Home Value Index", labels = scales::comma)


```
```{r}
ggplot(data=WakeCountySales, aes(x=YR, y=ZHVI, color = RegionName)) +
  geom_col() +
  labs(title = "Wake County Home Sales",
       x = "Year",
       y = "Home Value Index")+
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Home Value Index", labels = scales::comma)
```

a. What have been the overall trends in Wake County Home Values?
Increasing from 1996 to the 2008 recession, decreasing till about 2012 and then a continuous upward trend.
b. There were dips in home values in the past 20 years. What years did these occur?
1998, 2008
c. Based on the analysis, where would be the least expensive area to purchase home? Most expensive
area?
Zebulon/Wendell look to be the least expensive areas to purchase a home. Cary and Apex look like the most expensive areas to purchase homes. 
d. Are any area home values trending down? Is there one area that stands out compared to others?
No home values tend to be trending down. Zebulon has experienced very minimal growth over the years. New Hill in the scatterplot stands out as its home values drastically increased in 2018 being one of the most expensive home values.

### NC Rental Market
```{r}
Rentals <- select(SingleFamilyResidenceRental, RegionName, State, "2010-11", "2011-11","2012-11","2013-11","2014-11","2015-11","2016-11","2017-11","2018-10")
Rentals <- filter(Rentals, State == "NC")
Rentals <- filter(Rentals, RegionName %in% c("Asheville","Charlotte","Durham","Fayetteville","Raleigh", "Wilmington"))
Rentals <- rename(Rentals, "2010"="2010-11", "2011"="2011-11", "2012"="2012-11", "2013"="2013-11","2014"="2014-11", "2015"="2015-11","2016"="2016-11","2017"="2017-11","2018"="2018-10")
Rentals<-pivot_longer(Rentals, c("2010","2011","2012","2013","2014","2015","2016","2017","2018"), names_to = 'YR',values_to = 'ZHVI')
```

```{r}
ggplot(data=Rentals, aes(x=YR, y=ZHVI, color = RegionName)) +
  geom_point() +
  labs(title = "Rental Costs in NC Cities",
       x = "Year",
       y = "Average Rent")+
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Average Rent", labels = scales::comma)

```
```{r}
ggplot(data=Rentals, aes(x=YR, y=ZHVI, color = RegionName)) +
  geom_col() +
  labs(title = "Rental Costs in NC Cities",
       x = "Year",
       y = "Average Rent")+
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Average Rent", labels = scales::comma)
```

a. What has been the overall trend in the rental market around the state? Are there any cities that
have not followed this trend?
Most regions have seen a linear rise in rent prices. Fayetteville has not followed this trend, increasing and decreasing each year.
b. Where is the most expensive city to rent in? Least expensive?
Most expensive, Asheville. Least Expensive, Fayetteville.
c. You are trying decide between Wilmington and Asheville. Which market has the lowest rent?
Wilmington.

### Home Values in Select Rental Markets
```{r}
NCHomeSales <- select(SingleFamilyResidenceSales, RegionName, State, Metro, "1996-05", "1997-05", "1998-05","1999-05", "2000-05", "2001-05", "2002-05","2003-05", "2004-05", "2005-05","2006-05","2007-05", "2008-05","2009-05","2010-05", "2011-05","2012-05","2013-05","2014-05", "2015-05","2016-05","2017-05","2018-05")
NCHomeSales <-rename(NCHomeSales,"1996"="1996-05","1997"="1997-05","1998"="1998-05","1999"="1999-05","2000"="2000-05", "2001"="2001-05","2002"="2002-05","2003"="2003-05","2004"="2004-05","2005"="2005-05","2006"="2006-05","2007" ="2007-05","2008"= "2008-05","2009"= "2009-05","2010"= "2010-05","2011"="2011-05","2012" = "2012-05","2013"= "2013-05","2014"="2014-05", "2015"="2015-05","2016"= "2016-05","2017"="2017-05","2018"="2018-05")
NCHomeSales <- filter(NCHomeSales, State == "NC", Metro %in% c("Asheville", "Charlotte-Concord-Gastonia", "Raleigh", "Wilmington"))
NCHomeSales <- NCHomeSales %>% group_by(Metro)
NCHomeSales<-pivot_longer(NCHomeSales, c("1996","1997","1998","1999","2000","2001","2002","2003","2004","2005","2006","2007", "2008","2009","2010","2011","2012","2013","2014","2015","2016","2017","2018"), names_to = 'YR',values_to = 'ZHVI')
```

```{r}
ggplot(NCHomeSales, aes(x = Metro, y = ZHVI)) +
  geom_violin()+
  geom_boxplot(width=0.1)
```
a. According to the results, which market has the lowest median price (represented as horizontal bar in
box plot)? Charlotte-Concord-Gastonia
b. The violin plot will show density meaning the wider the plot is, the more observations occur within
that area. Which market has the most density around the median value of homes?
Charlotte-Concord-Gastonia

### Relocation Home Value Comparison
```{r}
NationalHomeSales <-select(SingleFamilyResidenceSales, RegionName, State, Metro, "1996-05", "1997-05", "1998-05","1999-05", "2000-05", "2001-05", "2002-05","2003-05", "2004-05", "2005-05","2006-05","2007-05", "2008-05","2009-05","2010-05", "2011-05","2012-05","2013-05","2014-05", "2015-05","2016-05","2017-05","2018-05")
NationalHomeSales <-rename(NationalHomeSales,"1996"="1996-05","1997"="1997-05","1998"="1998-05","1999"="1999-05","2000"="2000-05", "2001"="2001-05","2002"="2002-05","2003"="2003-05","2004"="2004-05","2005"="2005-05","2006"="2006-05","2007" ="2007-05","2008"= "2008-05","2009"= "2009-05","2010"= "2010-05","2011"="2011-05","2012" = "2012-05","2013"= "2013-05","2014"="2014-05", "2015"="2015-05","2016"= "2016-05","2017"="2017-05","2018"="2018-05")
NationalHomeSales <- filter(NationalHomeSales, State %in% c("CO","TX","NY","IL"), RegionName %in% c("Denver", "Houston","New York","Chicago"))
NationalHomeSales<-pivot_longer(NationalHomeSales, c("1996","1997","1998","1999","2000","2001","2002","2003","2004","2005","2006","2007", "2008","2009","2010","2011","2012","2013","2014","2015","2016","2017","2018"), names_to = 'YR',values_to = 'ZHVI')
NationalHomeSales <- NationalHomeSales %>% group_by(Metro)
```

```{r}
ggplot(data=NationalHomeSales, aes(x=YR, y=ZHVI)) +
  geom_point()+
  facet_wrap(~Metro)+
  labs(title = "City Home Value Comparison",
       x = "Year",
       y = "Home Index Value")+
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Home Index Value", labels = scales::comma)

```

a. Based on your analysis, which city’s housing is most affordable? Least affordable?
Houston is the most affordable. New York is the least affordable.
b. Which cities saw the largest change in prices over the past 5 years? Which city has remained more
consistent (i.e., no huge swings up or down in home values)?
New York and Denver have seen the largest change in the last five years with an increase of nearly 200,000. From the looks of it, Chicago has stayed the most consistent. I would consider Houston to also be quite consistent, but looks like they have seen a somewhat larger increase in the five year time period.
c. During the market downturn in 2012, which cities were most impacted? Which cities have
recovered?
New York and Chicago were most impacted. New York has recovered significantly, Chicago is slowly increasing since 2012 but not at the rate New York did. Denver was not impacted harshly in the 2012 downturn, but has seen significant growth since. 